Metabolites reflect variability introduced by mesoscale eddies in the North Pacific Subtropical Gyre

### 2024-11-14 manuscript/response_to_reviewers.txt

> Reviewer 1 | 04 Sep 2024 | 01:18
> #1
> This study examines the impact of mesoscale eddies on the metabolome and composition 
> of particulate organic matter in the North Pacific Subtropical Gyre. It fills a gap 
> in previous research by exploring the molecular composition of particulate organic 
> matter. The use both of targeted and untargeted metabolomics mass spectrometry to 
> detect molecules sensitive to environmental perturbation. The study finds that many 
> metabolites track biomass trends. Metabolites linked to eukaryotic phytoplankton are 
> enriched at the deep chlorophyll maximum of cyclonic features, while degradation 
> products are enriched in anticyclones. Additionally, depth variability and vertical 
> displacement of isopycnal layer constrain the importance of mesoscale eddies in 
> shaping particulate matter composition. Based on these results, we can infer the 
> biological communities and biogeochemical processes driven by mesoscale eddies at 
> both spatial and temporal scales. Viewing this complex dynamic process from the 
> chemical molecular scale is a perspective that sets this article apart from other 
> traditional methods in examining the changes in this system.
This is an accurate assessment of the manuscript.

> However, I still have some questions.
> Currently, the process of the entire system seems to be predictable. Although the 
> author has introduced and discussed some gaps, especially from the perspective of 
> chemical composition, and this article provides detailed data and reliable analysis, 
> it is not prominent enough from the perspective of the entire system. Perhaps the 
> transformation from space to time and the use of chemical composition to observe the 
> aging process of vortices is a great highlight. It is very difficult to focus on 
> scale eddies and the ecosystem evolution driven by them using traditional methods. 
> If that's the case, the structure and writing style of the paper may need to be 
> adjusted.
> This article focuses on the differences in chemical composition to understand this 
> process, but I always want to connect it with specific biological processes, such 
> as structural changes in biological communities, primary production rates, 
> zooplankton grazing rates, or bacterial production and respiration rates etc. 
> The author also mentioned these. I don't know if the author can add more relevant 
> evidence or discussion to allow readers to connect with these specific biological 
> groups and/or processes. I know it's difficult to integrate these data together, 
> but relevant discussions are possible.
> We already know that mesoscale eddies are complex and dynamic multidimensional 
> processes. The three-dimensional space combined with the temporal evolution of 
> ecosystem structure and function. The changes in the density surface driven by 
> physics during this process lead to the coupling reaction of physics chemistry 
> biology, which is the main thread. In fact, many other physical biological coupling 
> processes are similar, such as upwelling, internal waves/tides. Using the author's 
> method, I think similar situations can also be seen in such processes.
We have expanded the Discussion to touch on the concerns raised by this reviewer.
Specifically, we have mentioned possible links between the increased sulfonium
compounds in the eddy and DMS production, a hypothesis about the reason why
increased cyclonic biomass was not matched by increased export, and extensive
references to additional manuscripts with relevant data in lines 651-655 and 
689-718.

@@

> Specific suggestions,
> Table 1, the format is not appropriate
Table 1 has been removed entirely.

> Figures 4, and 6 can be placed together.
While we agree that the data presented in Figures 4 and 6 can be placed side-by-side
for ease of comparison, we believe that keeping these figures in their separate
locations within the manuscript allows for better narrative flow (since the HEE
results are not discussed until several sections later) and graphical design (as
proper visualization of NMDS requires that the x- and y- axes share units).

> Reviewer 2 | 04 Oct 2024 | 12:20
> #1
> Kumler et al. report in their manuscript “Metabolites reflect variability introduced 
> by mesoscale eddies in the North Pacific Subtropical Gyre” on a very impressive 
> dataset, investigating the polar part of the meta-metabolome of a > 0.2 µm fraction 
> of the plankton. They investigate samples taken at the Meso-scope and HEE experiment 
> and discuss their data in light of the eddies found during sampling.
> It is clearly a high-quality study that opens new perspectives. Publication after 
> revision is highly recommended.
This is an accurate assessment of the manuscript.

> Sampling:
> The metabolome at different depths can be strongly affected by the sampling time. 
> Day / night variations are substantial, mainly due to diel vertical migration but 
> also to photosynthetic state of the phytoplankton. Discussing the results sampling 
> times during the transect and during sampling of the HEE should be specified and 
> statistical analysis on the impact (or lack of impact) on the data should be provided.
We have added the time of day as an explanatory factor to our analysis and reported the
new values from the PERMANOVA in the text (Lines 220-226, 262-266 and 505-509) and figures, 
as well as adding a supplemental
figure to visually represent some of the variance due to time of day. We do see 
diel differences at the surface and DCM during the eddy transect and these have been
reported, with the 175 meter differences more likely to be due to chance and very little
diel difference in the HEE samples at all. We have additionally added a supplemental
figure identifying a few select compounds for which diel effects are clearly visible
(trehalose and sucrose) alongside other highly abundant compounds which show no diel
effects.

> Analytics:
> The high-quality analytics is appreciated. It should be verified if the variations 
> between biological replicates could be caused by the order of measurements in the 
> sequence. If >100 samples of 15 min are run in a row there could be a drift that would 
> be hidden because of the randomization. It would be easy to have a look at the outliers 
> (e.g. Fig. 2 lower panels) are measured early / late during the sequence.
We have double-checked that our normalization to internal standards removed any
analytical drift and can confirm that while significant drift occurred over
the course of the eddy transect sample run, the normalization removed this
effect and is not a major factor in the outliers visible in Fig. 2. (Lines 190-192)

> Blanks are mentioned but it is not shown how these were used for the data analysis.
We have added a note about how we used the methodological blanks to visually assess 
whether sample peak areas were above the background signal. (Lines 184-186)

### 2024-11-14 manuscript/manuscript.Rmd
Forgot to add the GPC fig to the manuscript
Adding spaces between 6AM and 6PM per Wei's recommendation

The HEE data was characterized by high inter-replicate variability relative to the samples from the MESO-SCOPE cruise, with 25 meter samples in particular highly variable in multivariate space (Figure 6). Despite this, we still saw the importance of sea level anomaly as a significant explanatory factor in the dataset (PERMANOVA R$^2_{SLA} = 0.090$, p-value = 0.017) as well as the larger depth differences between the surface (25 meters) and DCM (110 - 120 meters) (PERMANOVA R$^2_{depth} = 0.251$, p-value << 0.001). We saw no difference between the 6AM and 6PM samples (PERMANOVA R$^2_{time} = 0.031$, p-value = 0.367) despite a few known compounds (trehalose and sucrose) showing large differences, likely due to a majority of the mass features demonstrating no diel effect (Supplemental Figure 5).

@@

In all of the datasets analyzed here, we detected a significant difference in the composition of the metabolome between the adjacent eddies. This effect was strongest in the samples taken during the Lagrangian stations at the center of each eddy in the 2017 MESO-SCOPE cruise, with nearly a quarter of the total variance explained by the eddy from which the samples were taken. In the samples taken along the eddy transect, the largest effect was detected in the deepest samples taken from 175 meters and from the DCM, with much less of a response at 25 meters. This lack of SLA effect at the surface is unsurprising given the similarity in the biogeochemistry of the well-lit upper ocean previously reported [@Barone2022]. In the HEE samples, however, the opposite was true with a surprisingly larger SLA response at 25 meters than that at the DCM. In each case, the differences introduced by SLA were smaller than those of depth when compared directly, with even the high-resolution sampling around the DCM finding slightly more variance explained by the 40 meter difference in sampling depth than the 40 centimeter difference in SLA. This result agrees well with prior research showing the large effect of sampling depth on the metabolome [@Heal2021; @Kumler2023; @Bent2024]. Time of day was included as a significant factor in the MESO-SCOPE transect, with trehalose and sucrose in particular showing large differences between 6AM and 6PM as has been previously noted [@Muratore2022], though this effect was diminished at 175 meters.

@@

![](supplement/comb_gpc_patch.png)

*Supplemental Figure 4: Evidence against phosphorus limitation across the MESO-SCOPE eddy transect. Top row of plots (A) correspond to extracted ion chromatograms for 5 metabolites useful for estimating the phosphorus stress of an environment, colored by the depth from which the sample was taken. Middle row of plots shows the integrated peak area for each metabolite across the 11 stations of the transect with sub-panels for each depth independently, with each point colored by the corrected sea level anomaly (SLA). Bottom row of plots shows the ratio of choline to phosphocholine (left) and choline to glycerophosphocholine (right) as a function of corrected SLA with no statistically significant trend visible at any of the three depths.*

>>>>

The HEE data was characterized by high inter-replicate variability relative to the samples from the MESO-SCOPE cruise, with 25 meter samples in particular highly variable in multivariate space (Figure 6). Despite this, we still saw the importance of sea level anomaly as a significant explanatory factor in the dataset (PERMANOVA R$^2_{SLA} = 0.090$, p-value = 0.017) as well as the larger depth differences between the surface (25 meters) and DCM (110 - 120 meters) (PERMANOVA R$^2_{depth} = 0.251$, p-value << 0.001). We saw no difference between the 6 AM and 6 PM samples (PERMANOVA R$^2_{time} = 0.031$, p-value = 0.367) despite a few known compounds (trehalose and sucrose) showing large differences, likely due to a majority of the mass features demonstrating no diel effect (Supplemental Figure 5).

@@

In all of the datasets analyzed here, we detected a significant difference in the composition of the metabolome between the adjacent eddies. This effect was strongest in the samples taken during the Lagrangian stations at the center of each eddy in the 2017 MESO-SCOPE cruise, with nearly a quarter of the total variance explained by the eddy from which the samples were taken. In the samples taken along the eddy transect, the largest effect was detected in the deepest samples taken from 175 meters and from the DCM, with much less of a response at 25 meters. This lack of SLA effect at the surface is unsurprising given the similarity in the biogeochemistry of the well-lit upper ocean previously reported [@Barone2022]. In the HEE samples, however, the opposite was true with a surprisingly larger SLA response at 25 meters than that at the DCM. In each case, the differences introduced by SLA were smaller than those of depth when compared directly, with even the high-resolution sampling around the DCM finding slightly more variance explained by the 40 meter difference in sampling depth than the 40 centimeter difference in SLA. This result agrees well with prior research showing the large effect of sampling depth on the metabolome [@Heal2021; @Kumler2023; @Bent2024]. Time of day was included as a significant factor in the MESO-SCOPE transect, with trehalose and sucrose in particular showing large differences between 6 AM and 6 PM as has been previously noted [@Muratore2022], though this effect was diminished at 175 meters.

@@

![](supplement/comb_gpc_patch.png)
*Supplemental Figure 4: Evidence against phosphorus limitation across the MESO-SCOPE eddy transect. Top row of plots (A) correspond to extracted ion chromatograms for 5 metabolites useful for estimating the phosphorus stress of an environment, colored by the depth from which the sample was taken. Middle row of plots shows the integrated peak area for each metabolite across the 11 stations of the transect with sub-panels for each depth independently, with each point colored by the corrected sea level anomaly (SLA). Bottom row of plots shows the ratio of choline to phosphocholine (left) and choline to glycerophosphocholine (right) as a function of corrected SLA with no statistically significant trend visible at any of the three depths.*
